---
title: "Analysis of possible Labour targets"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Summary
There are **151** constituencies where the margin between the first and second parties is small and less than the combined share of other parties.

Out of these:

* **78** are constituencies where <span style="color: #C61148">Labour</span> came second
    + **51** are <span style="color:#0096DB">Conservative</span> constituencies where <span style="color: #C61148">Labour</span> came second
    + **25** are <span style="color:#FCCA46">SNP</span> constituencies where <span style="color: #C61148">Labour</span> came second
    + **2** are **Plaid Cymru** constituencies where <span style="color: #C61148">Labour</span> came second
    
There are **92** Conservative constituencies where the margin between the first and second parties is small and the second party is progressive (Labour, SNP, LD, Green).

```{r, echo=TRUE, message=FALSE, warning=F}
library(dplyr)
library(readr)
library(plotly)
library(DT)
ge_2017 <- read_csv("http://researchbriefings.files.parliament.uk/documents/CBP-7979/hocl-ge2017-results-full.csv")

```

```{r}
party_colors <- 
"Party,Color
Lab,#C61148
Con,#0096DB
SNP,#FCCA46
LD,#F37A48"
party_colors <- read.csv(textConnection(party_colors)) %>% mutate(Color=as.character(Color))
```

```{r, message=FALSE, warning=FALSE}
results <- ge_2017 %>%
    group_by(constituency_name) %>%
    top_n(n = 2, wt=share) %>%
    select(constituency_name, party_abbreviation, region_name, share) %>%
    filter(!(region_name %in% c("Northern Ireland"))) %>%
    mutate(margin=share-lead(share), second_party=lead(party_abbreviation), third_party_share=1-share-lead(share)) %>%
    filter(!is.na(margin)) %>%
    arrange(-margin) %>%
    left_join(party_colors, by=c("party_abbreviation"="Party")) %>%
    mutate(Color=ifelse(is.na(Color), "grey", Color), Color=as.factor(Color)) %>%
    mutate(margin=margin*100, share=share*100, third_party_share=third_party_share*100) %>%
    ungroup()
```
## Marginal constituencies where Labour is second
```{r}
possible_gains <- results %>%
    filter(second_party=="Lab" & margin < third_party_share & margin < 15)

results %>%
    filter(second_party=="Lab") %>%
    plot_ly() %>%
add_lines(x=list(0, 35), y=list(0, 35), line=list(width=0), fillcolor="rgba(10,10,10,0.1)", fill="tozeroy") %>%
add_lines(x=list(0, 35), y=list(15, 15), line=list(width=0), fillcolor="rgba(10,10,10,0.1)", fill="tozeroy") %>%
    add_markers(x=~third_party_share, mode="markers", y=~margin, marker=list(color=~Color),
            type="scatter", text=~paste("<b>",constituency_name, "</b>",
                                        "<br>",
                                        "First party: ", party_abbreviation,
                                        "<br>",
                                        "Second party: ", second_party)) %>%
           layout(xaxis=list(title="Small party combined share (%)"),
                 yaxis=list(title="Margin (%)"),
                 showlegend=F)
```

## Constituencies where the small parties have a share less than the margin and the margin is less than 15%

```{r, message=FALSE, warning=FALSE}
possible_gains %>%
    arrange(margin) %>%
    select(-Color) %>%
    mutate(share=paste0(format(round(share, 1), nsmall = 1), "%"),
          margin=paste0(format(round(margin, 1), nsmall = 1), "%"),
          third_party_share=paste0(format(round(third_party_share, 1), nsmall = 1), "%")) %>%
    datatable(options=list(pageLength=20),
              colnames=gsub("_", " ", gsub("(^|_)([[:alpha:]])", "\\1\\U\\2", colnames(.), perl=TRUE))) %>%
formatStyle(
  'party_abbreviation',
    target="row",
  backgroundColor = styleEqual(party_colors$Party, party_colors$Color)
) %>% formatStyle(
  'second_party',
  backgroundColor = styleEqual(party_colors$Party, party_colors$Color)
)
```

## Filter by Conservative constituencies

```{r, message=FALSE, warning=FALSE}
possible_gains %>%
    arrange(margin) %>%
    filter(party_abbreviation=="Con" & second_party=="Lab") %>%
    select(-Color) %>%
    mutate(share=paste0(format(round(share, 1), nsmall = 1), "%"),
          margin=paste0(format(round(margin, 1), nsmall = 1), "%"),
          third_party_share=paste0(format(round(third_party_share, 1), nsmall = 1), "%")) %>%
    datatable(options = list(pageLength=20),
              colnames=gsub("_", " ", gsub("(^|_)([[:alpha:]])", "\\1\\U\\2", colnames(.), perl=TRUE))) %>%
formatStyle(
  'party_abbreviation',
    target="row",
  backgroundColor = styleEqual(party_colors$Party, party_colors$Color)
) %>% formatStyle(
  'second_party',
  backgroundColor = styleEqual(party_colors$Party, party_colors$Color)
)
```

## Conservative constituencies where the second party is Labour, LD, SNP or Green

```{r, message=FALSE, warning=FALSE}
results %>%
    filter(party_abbreviation == "Con" & second_party %in% c("Lab", "LD", "Green", "SNP")) %>%
arrange(margin) %>%
    select(-Color) %>%
    mutate(share=paste0(format(round(share, 1), nsmall = 1), "%"),
          margin=paste0(format(round(margin, 1), nsmall = 1), "%"),
          third_party_share=paste0(format(round(third_party_share, 1), nsmall = 1), "%")) %>%
    datatable(options = list(pageLength=20),
              colnames=gsub("_", " ", gsub("(^|_)([[:alpha:]])", "\\1\\U\\2", colnames(.), perl=TRUE))) %>%
formatStyle(
  'party_abbreviation',
    target="row",
  backgroundColor = styleEqual(party_colors$Party, party_colors$Color)
) %>% formatStyle(
  'second_party',
  backgroundColor = styleEqual(party_colors$Party, party_colors$Color)
)
```

